---
title: "Week 6 Coding Lesson"
author: "Jordan Marisa Davis"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

## Part 2: A rank-based test example (Mann Whitney U)

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Is there evidence for a significant difference in ranks (medians) between the populations "gp_1" and "gp_2" were drawn?

```{r}
hist(gp_1)
```
```{r}
hist(gp_2)
```

If I want to compare ranks between gp_1 and gp_2, what are some reaosns I might choose a rank-based test?

1. Not clearly normally distributed from exploratory histograms
2. Somewhat small sample size (n = 15 for each)
3. I've decided that ranks (or, medians) are a more valueable metric to compare for these data.

Here's we'll perform Mann-Whiteney U to answer "Is tehre a significant difference in ranks(medians) between "gp_1" and "gp_2"? 

```{r}
my_mwu <- wilcox.test(gp_1, gp_2)
```

What does that p-value of 0.28 actually mean? It means that if the null hypothesis is true (these samples were drawn from populations with the same median), there is a probability of 0.28 that we could have found median values at least as different as ours by chance. In other words: not sufficient evidence to reject the null hypothesis of equal ranks (or medians) using a significance level of 0.05.

## Part 3: Simple Linear Regression

We’ll exploring the relationship between two continuous variables, using the penguins dataset from the palmerpenguins package. 

Here, we will explore the relationship between flipper length and body mass for penguins, including all 3 penguin species in the dataset. 

### A. Look at it!

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g))+
  geom_point()
```

We should ask questions about our exploratory visualization, like:

- Does it look like a linear relationship makes sense?
- Do we have any concerns about modeling as a linear relationship?
- Any notable outliers?
- Initial thoughts about homoscedasticity (explored more later)?

Here, it looks like overall a linear relationship between flipper length and body mass makes sense here (moving forward, we’re learn how to include species and sex as part of the model, but for now we’ll just use the single exploratory variable "flipper_length_mm" 

### B. Model it

Once we've decided that a lienar relationship makes sense, we'll model it using lm()

Note that we haven't checked all assumptions yet. That is because a lot of our assumptions for linear regression are based on model residuals (e.g. normality & homoscedasticity of residuals), which we can’t calculate until after we find the predicted values from the model.

```{r}
# Linear model, stored as penguins_lm
penguins_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

# Return the complete overview
summary(penguins_lm)
```

Some interpretations:

- Both the intercept and flipper_length_mm coefficients are significantly different from zero (not super interesting)

- The multiple R^2 value is 0.759, meaning that 75.9% of variance in body mass is explained by flipper length

### Part C. Access Model Outputs

We can access the coefficients for the model using:

 - The slope is 49.69 (g/mm)
 - The y-intercept is ~5780.83 (g)
 - The full equation is mass = 49.69*(flipper length) + (~5780.83)
 
 We an use the "broom::tidy()" function to get the model outputs in nice data frame format:
```{r}
penguin_lm_tidy <- broom::tidy(penguins_lm)
```
 
```{r}
# Get the intercept:
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int
```
```{r}
# Then to get the flipper_length coefficient:
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef
```
```{r}
# Metrics at a glance:
penguin_lm_out <- broom::glance(penguins_lm)
penguin_lm_out
```

We can use the results of both to write a statement about the model that will automatically update if anything about the model changes! Make sure to look at the .Rmd (not just this knitted html) to learn how to reference the outputs automatically in text. For example:

“Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found (β = 49.686, F(1,340) = 1070.7, p < 0.001) with an R2 of 0.759.”

Note: This might seem really tedious to write out, but the advantages are worth it. All values will be automatically updated when the model is updated! Reproducible and way less opportunity for human error. Plus, once you have this template statement made, you can reuse it for future regression models and just replace "penguins_lm_out" and "penguins_coef" with the appropriate objects for your new model!

Note that I use “p < 0.001” here if the p-value is very small - this is somewhat standard.

### Part D. Explore Model Assumptions 

Recall that we have assumptions for lienar regression we need to explore, some related to the residuals. 

- Linearly related varaibles (CHECK - already looked & thoght hard)
- Normally distributed residuals 
- Homoscedasticity (constant residuals variance)
- iid residuals (no serial correlation) - more often a concern in time series data 

Use the "plot()" function on the model, which will automatically create four useful visualizations to consider assumptions.

```{r}
plot(penguins_lm)
```
Notice that four plots show up. What do they show? Make sure to watch Part 2 of the lecture, which discusses how we can interpret each of these diagnostic plots.

 - The first one: fitted values vs. residuals
 - The second one: QQ-plot for residuals 
 - The third one: another way of looking at fitted vs. residuals (there are just standardized residuals, but you can interpret it the same way)
 - The fourth one: Cook's distance, a measure of "influence" or "levereage" that individual points have on the model - often considered a way to explore outliers 
 
See the Week 6 Part 2 Lecture video for more information about how to interpret these outcomes, but in summary: graphs 1 & 3 are useful for thinking about homoscedasticity; graph 2 (QQ plot) helps us consider normality of residuals; graph 4 reveals the Cook’s distance (a measure of how much leverage any single observation has on the model).

### E. Visualize the model

Now that we’ve explore the assumptions and have decided that linear regression is a valid tool to describe the relationship between flipper length and body mass, let’s look at the model.

- Use "geom_smooth(method = "lm") to add a linear model to an existing scatterplot 
- Use "stat_cor()" and/or "stat_regline_equation()" to add equation information directly to the plot panel, at an x-and y position that you specify 

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)
```

### F. Find Pearson's r for correlation:

In lecture we talked about the coefficient of determination, R2, which tells us how much of the variance in the dependent variable is explained by the model.

We might also want to explore the strength of the correlation (degree of relationship) between two variables which, for two linearly related continuous variables, can be expressed using Pearson’s r.

Pearson’s r ranges in value from -1 (perfectly negatively correlated - as one variable increases the other decreases) to 1 (perfectly positively correlated - as one variable increases the other increases). A correlation of 0 means that there is no degree of relationship between the two variables.

Typical guidelines look something like this (there’s wiggle room in there):

- r = 0: no correlation
- r < |0.3|: weak correlation
- r between |0.3| and |0.7|: moderate correlation
- r > |0.7|: strong correlation 

We'll use "cor.test()" function, adding the two vector ("flipper_length_mm" and "body_mass_g") as the arguments. The function reports the Pearson’s r value, and performs a hypothesis test with null hypothesis that the correlation = 0.

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)
```

Here, we see that there is a strong positive correlation between penguin flipper length and body mass (r = 0.87, t(340) = 32.72, p < 0.001).